---
title: Palisades Wildfire Burned Acres Forcast
author: "Haiyang Wu"
date: "May 8, 2025"
output:
  pdf_document: null
  highlight: default
  keep_tex: no
  fig_caption: yes
  latex_engine: pdflatex
  html_document:
    df_print: paged
affiliation: Department of Statistics, UConn
fontsize: 11pt
geometry: margin=1in
---

```{r, include = FALSE}
library(car)
library(MASS) 
```

# 1 Introduction  {#sec:intro}
Los Angeles County experiences destructive wildfires on a yearly basis.  Incident
command posts issue situation reports throughout an event, logging **(i)** the
latest containment percentage, ** (ii)** personnel deployed, and **(iii)** the
cumulative acreage burned.  The *CleanData.csv* file supplied for this project
contains **344 timestamped status rows** covering the **\~5‑week window between
7 January 2025 and 10 February 2025** for a single large incident.

> **Research question** *Can we predict the eventual* **Acres Burned** *from the
> information available mid‑incident—specifically containment percentage and
> staffing levels—so that managers can estimate how much additional land is at
> risk?*

A reliable regression model would allow operations chiefs to allocate crews
and equipment more efficiently during the active phase of a fire rather than
waiting for post‑incident totals.


# 2 Methodology  {#sec:methods}

## 2.1 Data preparation  {#sec:data}
```{r data-load}
# raw import (base R)
wild <- read.csv("CleanData.csv", stringsAsFactors = FALSE)

# basic sanity‑check
cat("Rows before NA drop:", nrow(wild), "\n")

# omit records with any missing core variable
wild <- na.omit(wild[ , c("Update.Timestamp", "Acres.Burned..Size.",
                         "Containment.Percent", "Total.Personnel")])
cat("Rows after NA drop:", nrow(wild), "\n")

# convert timestamp to POSIXct and extract calendar features
wild$UpdateTime <- strptime(wild$Update.Timestamp,
                            format = "%m/%d/%Y %I:%M %p")
wild$Date       <- as.Date(wild$UpdateTime)
```

```{r eda, fig.cap="Exploratory plots of the key variables."}
par(mfrow = c(1,3))
hist(wild$Acres.Burned..Size., 30, main = "Acres Burned", xlab = "Acres")
hist(wild$Containment.Percent, 30, main = "Containment %", xlab = "%")
hist(wild$Total.Personnel, 30, main = "Personnel", xlab = "Number")
par(mfrow = c(1,1))

pairs(wild[, c("Acres.Burned..Size.",
               "Containment.Percent",
               "Total.Personnel")],
      main = "Scatter‑plot matrix")
```

> *Observation.* Acres burned is right‑skewed (long tail at small fires), while
> containment is heavily left‑skewed (many early‑stage updates).  Personnel and
> containment appear weakly correlated, suggesting that collinearity is not a
> major concern.

